---
title: "Brownlows Bets"
author: "Deniss Gonzalez Ramirez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
source(file="deniss.R")
```


<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.

## Our strategy is to build a Function and the use it to answer the three questions. 

The function will be given a class_size by number of students, odds, and Number_of_times brownlow made the bet.  It will give you back how much Brownlows earnings were. We will use it to answer our three questions below. If you can, improve this function by giving default values to some of its inputs. 

# In the class and payoff as described  If you highered 40 brownlow substitute teachers to do the same thing that Brownlow did then.
## what is their average earnings
## the standard deviation and a 95% confidence interval for their earnings?
  
<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows winnings (or losings)-->

The code block below explains that the array has 1000 elements inside of it. Each element of the array Data1 represents how much a brownlow gave 2:1 odds and made 637 bets made as a profit. Had this been done 1000 times is what is depicted below. The data also had to be sorted from low to high in order to determine the 95% confidence level and to determine what odds were the most profitable. The middle of it runs from 25 and 975.

```{r}
Data1<-array(dim = 1000)
Data1<-as.numeric(Data1)
for(i in 1:1000){
  Data1[i] = deniss()
}
sdata1<-sort(Data1)
hist(sdata1)
abline(v=sdata1[25],col="red")
abline(v=sdata1[975],col="blue")
low<-sdata1[25]
high<-sdata1[975]

```

We determined that 40 brownlows was not enough so the test was configured to have 1000 elements. This would provide sufficient evidence
The 95% confidence interval is `r low` to `r high`

# What are the fair odds? Use manipulation of odds in the function to tune Brownlows earnings to zero. 


<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 637 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->

After tinkering with the function at hand we determined that 4.39 are fair odds.

```{r}
Data2<-array(dim = 200)
Data2<-as.numeric(Data2)
for(i in 1:200){
  Data2[i] = deniss(odds = 4.39)
}
sdata2<-sort(Data2)
hist(sdata2)
abline(v=sdata2[5],col="red")
abline(v=sdata2[195],col="blue")
low<-sdata2[5]
high<-sdata2[195]

```

The plot above seems to show that zero is centered at the odds of roughly 4.39. This indicates the odds that Brownlow would have to give in order for his profits to be close to zero.

# What size of class makes the 2:1 bet fair?  Use manipulation of number of Students to tune Brownlows earnings to zero.

<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting profits!-->

```{r}
Data3<-array(dim = 200)
Data3<-as.numeric(Data3)
for(i in 1:200){
  Data3[i] = deniss(numClass = 29)
}
sdata3<-sort(Data3)
hist(sdata3)
abline(v=sdata3[5],col="red")
abline(v=sdata3[195],col="blue")
low<-sdata3[5]
high<-sdata3[195]

```

The histogram above shows that in order to make the 2:1 bets fair then the number of students in a given class should be about 29. By trying different numbers narrowing down the number of people in the class. We were able to deduce that the less people there was in the class, the less money Brownlow made from bets.

The red and blue line in the plot above represents the 95% confidence interval of his earnings.



# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->

Based on the information above, we can conclude that professor Brownlow would have to operate at the odds of 4.39 in order for his profits to center on zero. In order to obtain this result, His class should be at 29 students in order for his current scheme to be at zero profit. We were able to obtain this number by narrowing the number down from 35 and testing which would be closer to zero.
